A star formation event in the leading arm of the Magellanic stream triggered by passage through the outer Galactic disk

### 2018-07-18 ms.tex
edits

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data release 2 (\DR{2}; \citealt{Gaia-Collaboration:2018, Lindegren:2018}) cross-matched to the \pans\ data release 1 photometry ($griz$ magnitudes; \citealt{Chambers:2016}) to search for distant, comoving multiplets of blue stars.\footnote{We use the cross-match provided by the \gaia\ science archive at \url{https://gea.esac.esa.int/archive/}.}
We initially query all stars from the \gaia--\pans\ cross-match that have $(g - i) < 1$, parallax $\varpi < 1~\textrm{mas}$, and Galactic latitude $|b| > 10^\circ$.

@@

\vspace{5mm}
\facilities{TODO}

>>>>

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data release 2 (\DR{2}; \citealt{Gaia-Collaboration:2018, Lindegren:2018}) to search for distant, comoving multiplets of blue stars.
We initially select all stars from \gaia\ with $\textrm{parallax} < 1$, $-0.5 < \bprp < 0$, $G < 20$, and Galactic latitude $|b| > 20^\circ$ (see \sectionname~\ref{sec:queries} for database query).
We further exclude stars within a $15^\circ$ radius from the LMC, and a $8^\circ$ radius from the SMC --- this leaves just 27,895 stars.
We then cross-match this catalog to itself with both sky positions and proper motions: we search for pairs of stars that have separations $s < 0.5^\circ$ and proper motion differences $|\Delta \mu| < 0.5~\masyr$.
We then combine mutually-connected comoving pairs into small groups of stars that are colocated on the sky and comoving in proper motions; We filter out groups that have $<4$ members.
We cross-match the mean sky positions of the groups to locations of local group galaxies \todo{citation} \citep{Mcconaghie} and Milky Way globular clusters \todo{citation} \citep{Harris} and filter out all groups that lie within 1 degree of these known objects.
After these filters, just one group of blue, comoving stars remains at $(\textrm{RA}, \textrm{Dec}) \sim (179, -29)^\circ$.

We then query all objects from the \gaia\ \DR{2} catalog within a polygon centered on the nominal position of the group (see above), with a width of $12^\circ$ and a height of $8^\circ$ in the Magellanic stream coordinate system \citep{Nidever:2008} (see \sectionname~\ref{sec:queries} for the database query).
We cross-match the resulting catalog with \acronym{GALEX} \todo{citation} using a positional crossmatch requiring source separation $<2~\textrm{arcsec}$, and with \acronym{PS1} \todo{citation}\citep{Chambers:XX} using the cross-match provided by the \gaia\ science archive.

\todo{Left off here}

@@

\appendix

\section{Queries}
\label{sec:queries}

Initial query to select very blue stars away from the Galactic plane:
\begin{verbatim}
SELECT * FROM gaiadr2.gaia_source
WHERE parallax < 1
AND (bp_rp > -0.5) AND (bp_rp < 0)
AND phot_g_mean_mag < 20
AND ABS(b) > 20
\end{verbatim}

Query to retrieve \gaia\ data around the blue, comoving group found and discussed in \sectionname~\ref{sec:data}:
\begin{verbatim}
SELECT *
FROM gaiadr2.gaia_source
WHERE (parallax < 0.5 OR parallax IS NULL) AND
    CONTAINS(POINT('ICRS', ra, dec),
             POLYGON('ICRS',
                     182.1954909980203, -35.51467868323814,
                     172.7240896740689, -33.70157733692543,
                     176.3745742549848, -22.277118488646703,
                     184.81885184435987, -24.236022801996057)) = 1 AND
    (pmra > -10) AND (pmra < 10) AND
    (pmdec > -10) AND (pmdec < 10)
\end{verbatim}
